Predicting Kamala Harris's Victory in the 2024 US Election

### 2024-10-31 paper/paper.qmd
Update the Title, Abstract, Discussion

---
title: "Predicting Kamala Harris's Victory in the 2024 US Election"
subtitle: "An Investigation into Poll Data Using Multiple Linear Regression"
author: 
  - Xuanle Zhou
  - Yongqi Liu
  - Yuxuan Wei
thanks: "Code and data are available at: https://github.com/wyx827/2024USpresidentialelection.git"
date: today
date-format: long
toc: true
abstract: "The study predicting the 2024 U.S. presidential election is important as it provides the predicted outcomes for the US election. Analysis is based on state-by-state electoral votes, and these outcomes can influence campaign strategies and shape public discourse. The result indicates Kamala Harris will win by over 270 votes in this election cycle."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

```{r}
#| label: fig-end-date
#| fig-cap: Trend of Support Percentage for Donald Trump Over Days Until Election. Fluctuations in support levels are observed as the election approaches, with a slight downward trend indicated by the dashed curve.
#| echo: false
#| warning: false
#| message: false

ggplot(analysis_data, aes(x = days_until_election, y = percent)) + 
  geom_line(color = "skyblue", size = 1.2) +  
  geom_smooth(method = "loess", color = "pink", linetype = "dashed", size = 1) +  
  labs(x = "Days Until Election", y = "Support Percentage") + 
  theme_minimal(base_size = 14) +  
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),  #
    axis.title.x = element_text(margin = margin(t = 10)),  
    axis.title.y = element_text(margin = margin(r = 10)),  
    axis.text = element_text(color = "#555555"),  
    panel.grid.minor = element_blank()  )
```

@@

## Overview of Predictive Modeling for the 2024 U.S. Election
This paper employs a regression model to forecast the 2024 U.S. Presidential Election outcomes between Kamala Harris and Donald Trump. By analyzing historical state-level voting data, we estimated Trump’s vote percentages across states and calculated each candidate’s projected electoral votes. Our model predicts a significant victory for Harris, who is projected to secure 387 electoral votes compared to Trump’s 151. This finding underscores the impact of demographic and geographic factors on the U.S. electoral landscape. A geographic visualization further emphasizes the anticipated state-by-state outcomes and reveals enduring regional voting patterns.

@@

@@

## Weakness
Despite the model's strengths, it has limitations that requires further exploration. The assumption of linear relationships between predictors may oversimplify complex voter behaviors and social influences. The dataset simplifies the potentioanl influecne behind the US election. Additionally, the reliance on past data assumes that historical trends will persist, which may not hold true in the rapidly shifting political climate of the U.S. To address these issues, future models could benefit from incorporating voter turnout projections and real-time sentiment analysis, potentially drawing data from social media platforms and news sources to gauge public opinion. Moreover, a comparative post-election analysis could offer valuable insights, revealing areas where the model succeeded and where it may need refinement. This iterative approach would help create a more resilient and adaptable electoral forecasting model.

@@

The diverse sampling approach not only ensures a broad range of opinions is captured but also complements probability-based sampling, which accurately reflects the overall population. Furthermore, reweighting the data according to U.S. Census demographics strengthens the credibility of the results by ensuring demographic accuracy. Additionally, rotating questions and answer choices helps mitigate bias, further improving the reliability of the data. Finally, the use of online surveys offers a cost-effective solution for efficient data collection.\
* **Disadvantages:**

>>>>

---
title: "Predicting Kamala Harris's Victory in the 2024 US Election"
subtitle: "Poll Data Predicts Kamala Harris Win with 387 Electoral Votes"
author: 
  - Xuanle Zhou
  - Yongqi Liu
  - Yuxuan Wei
thanks: "Code and data are available at: https://github.com/wyx827/2024USpresidentialelection.git"
date: today
date-format: long
toc: true
abstract: "This study provides a prediction for the 2024 U.S. presidential election, presenting potential outcomes based on a state-by-state electoral vote analysis. These projections can guide campaign strategies and influence public discourse. The results suggest Kamala Harris will secure a win with 387 electoral votes, significantly surpassing the 270 needed for victory. In contrast, Donald Trump is predicted to receive 151 electoral votes."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

```{r}
#| label: fig-end-date
#| fig-cap: Trend of Support Percentage for Donald Trump Over Days Until Election. Fluctuations in support levels are observed as the election approaches, with a slight downward trend indicated by the dashed curve.
#| echo: false
#| warning: false
#| message: false
ggplot(analysis_data, aes(x = days_until_election, y = percent)) + 
    geom_line(color = "skyblue", size = 1) + 
    geom_smooth(method = "loess", color = "pink", linetype = "dashed", size = 1) + 
    labs(x = "Days Until Election", y = "Support Percentage") + 
    theme_minimal(base_size = 14) + 
    scale_x_reverse() + 
    theme(
        plot.title = element_text(face = "bold", size = 16, hjust = 0.5),  
        axis.title.x = element_text(margin = margin(t = 10)), 
        axis.title.y = element_text(margin = margin(r = 10)), 
        axis.text = element_text(color = "#555555"), 
        panel.grid.minor = element_blank()
    )

@@

## Overview of Predictive Modeling for the 2024 U.S. Election
This paper forecasts the 2024 U.S. Presidential Election outcomes between Kamala Harris and Donald Trump, presenting a predicted electoral college outcome based on a fitted model. By analyzing poll data and historical state-level voting patterns, we estimated Trump’s vote percentages across states and calculated projected electoral votes for each candidate. Our model anticipates a significant victory for Harris, who is projected to secure 387 electoral votes compared to Trump’s 151. This finding underscores the impact of demographic and geographic factors on the U.S. electoral landscape.

A geographic visualization further emphasizes the anticipated state-by-state outcomes and reveals enduring regional voting patterns. Key predictors, such as state-level variations and transparency scores, strongly influence polling outcomes for both Donald Trump and Kamala Harris. While our model projects a potential Harris victory with a majority of electoral votes, challenges inherent in electoral predictions remain. We acknowledge that regional variations, particularly in battleground states like Florida and Arizona, introduce uncertainty that may affect the final outcome.

@@

## Weakness
Despite the model's strengths, it has limitations, which present a critical challenge and requires further exploration. The assumption of linear relationships between predictors may oversimplify complex voter behaviors and social influences. The dataset simplifies the potentioanl influecne behind the US election.

Moreover, polling averages, though valuable, are subject to biases that may skew results. For example, factors like nonresponse bias—where certain demographic groups, such as less civically engaged individuals, are less likely to participate—may influence results more heavily in this election cycle. Additionally, reliance on past election data introduces a risk of recency bias, as seen in public expectations shaped by Trump’s unexpected win in 2016 and close finish in 2020. Yet, as Nate Silver notes, intuition or prior election outcomes can be misleading, given the unique dynamics of each election. Polling errors, such as those seen in 2016 and 2020, further emphasize the need for cautious interpretation of our model’s findings.

@silver2024gut also points out that the voters for Trump are less likely to finish the polls because of “Shy Voters”. This may lead to the bias in the survey sample collection. The deficiency of education background and sample credibility both influence the prediction.

Additionally, due to limited data available for predicting outcomes in every state, some state projections rely on historical voting patterns, with the assumption that these trends will persist. However, this approach may be challenged by the rapidly changing political landscape in the U.S., where evolving demographics and shifting public sentiment could lead to unexpected results.

@@

@@

The diverse sampling approach not only ensures a broad range of opinions is captured but also complements probability-based sampling, which accurately reflects the overall population. Furthermore, reweighting the data according to U.S. Census demographics strengthens the credibility of the results by ensuring demographic accuracy. Additionally, rotating questions and answer choices helps mitigate bias, further improving the reliability of the data. Finally, the use of online surveys offers a cost-effective solution for efficient data collection.\

* **Disadvantages:**
